---
title: "20181019m Matrices"
author: "cheungngo"
date: "19 October 2018"
output: word_document
---

### Definition of matrices

```{r}
A = matrix(c(1,3,2,2,8,9), byrow = T, nrow = 2); A
```

### Multiplying a matrix with a number

```{r}
A*3
```

### Transpose of matrix

```{r}
t(A)
```

### Sum of matrices

```{r}
A = matrix(c(1,2,2,1),ncol = 2); A
B = matrix(c(3,1,4,2),ncol = 2); B
A + B
```

### Multiplication of matrices

By row from left to right, row(i) of first matrix * col(i) of second matrix  
i.e. number of row depends on the row of first matrix  
and number of col depends on the col of the second matrix  

a 2x3 %*% 3x2 matrix would thereby give a 2x2 matrix;  
while a 3x2 %*% 2x3 matrix would give a 3x3 matrix

```{r}
A = matrix(c(1,3,2,2,3,5),ncol = 3); A
B = matrix(c(2,0,1,1,-1,2),ncol = 2); B
C = t(A); C
D = t(B); D
```

```{r}
A %*% B
C %*% D
```

### Special matrices

```{r}
# Diagonal matrix
E = diag(c(1,2,3)); E
A; diag(A)
```

Square matrix: n x n matrix  
Symmetric matrix: A = t(A)  
0-matrix and 1-matrix (Note 1-matrix is denoted as J)  
Identity matrix: 1 on the diagnoal; also note IA = AI = A  

### Inverse of matrices

For a matrix A, a matrix B is said to be an inverse matrix if:  
AB = BA = I  
(A must be a square matrix)
(not all square matrix has an inverse)

```{r}
A = matrix(c(1,2,3,4),ncol = 2); A
B = solve(A); B
A %*% B
B %*% A
```

