---
title: "20181012a Ryacas basics"
author: "cheungngo"
date: "October 12, 2018"
output: word_document
---

```{r, message=FALSE}
library(Ryacas)
```

## Common functions in Ryacas

###

### yacas()

```{r}
a = yacas(expression(x^2+1))
b = expression(x^2+1)
class(a)
class(b)
```

###

### Eval()

```{r}
a = yacas(expression(x^2+1))
Eval(a,list(x=9))
```

```{r}
Eval(a,list(x=c(1:9)))
```

###

### yacas expressions

```{r}
# i.e. the quotation ''
# inside the quotation would be the expressions which is considered non-proper R expressions
d = yacas('D(x)Sin(x)')
print(d)
```

###

### Sym()

```{r}
x = Sym('x') # remember the quotation
x + 10 # operations on Sym Objects directly without yacas()
```

```{r}
# One can apply sin, cos, tan, deriv, integrate and other functions to Sym objects
deriv(2*x^2,x)
```

###

## Simple Yacas calculations

###

### Set() or n:=
#### For global variables

```{r}
yacas('n:=Cos(x)')
yacas('n')
yacas('n:=n+n')
yacas('Set(m,Sin(x))')
yacas('m')
yacas('m+m')
# Remerber the global variables cannot be printed
```

### Further application for Set()
#### assuming we would like to calculate area = pi * r^2

```{r}
pi = Sym('pi')
r = Sym('r')
# Suppose r = 2
Set(pi,3.14)
Set(r,2)
pi * r^2
```

###

### Clear()
#### Note that after clearing a variable (with both Set and Sym applied), the Sym would remain

```{r}
pi
Clear(pi)
pi
```

###

### N()
#### Returning the exact value
#### (otherwise it would remain a rational number)

```{r}
yacas('9/5')
yacas('N(9/5)')
```

```{r}
# precision in N()
yacas('N(Exp(1),5)')
```

###

### %
#### Results from the previous line

```{r}
yacas('(1+x)^3')
yacas('%')
k = Sym('%')
k
```

###

### PrettyForm()

```{r}
yacas('%')
yacas('PrettyForm(%)')
```

