---
title: "20181019n systems of linear equations"
author: "cheungngo"
date: "19 October 2018"
output: word_document
---

```{r,message=FALSE}
library(Ryacas)
```


### Solving systems of linear equations

x1 + x2 + x3 == 6  
2x1 + 4x2 + x3 == 5  
2x1 + 3x2 + x3 == 6  

The system can be re-written in matrix form: AX = B

```{r}
A = matrix(c(1,2,2,1,4,3,1,1,1),ncol = 3); A
x1 = Sym('x1'); x2 = Sym('x2'); x3 = Sym('x3')
X = matrix(c(x1,x2,x3),ncol = 1); X
B = matrix(c(6,5,6),ncol = 1); B
```

To solve X, we can have A-1B

```{r}
solve(A)%*%B
```

x1 = 2; x2 = -1; x3 = 5


### Elementary row operations

Without computer, we can also try to solve such linear equations by hand  

Consider again:  
x1 + x2 + x3 == 6  
2x1 + 4x2 + x3 == 5  
2x1 + 3x2 + x3 == 6  

```{r}
Aug = matrix(c(1,2,2,1,4,3,1,1,1,6,5,6),ncol = 4); Aug
```

We aim at creating a echelon form, i.e. 1 on the diag and 0 on the lower tri

```{r}
# To create 0 in the second row:
# R2 = R2 - 2*R1
Aug[2,] = Aug[2,]-2*Aug[1,]; Aug
```

```{r}
# To create 1 in the second row:
# R2 = R2 / 2
Aug[2,] = Aug[2,] / 2; Aug
```

```{r}
# To create 0 in the third row:
# R3 = R3 - 2*R1
Aug[3,] = Aug[3,] - 2*Aug[1,]; Aug
```

```{r}
# To create another 0 in the third row:
# R3 = R3 - R2
Aug[3,] = Aug[3,] - Aug[2,]; Aug
```

```{r}
# To create 1 in the third row:
# R3 = R3 / R2
Aug[3,] = Aug[3,] * -2; Aug
```

Hence, we can get:  
x3 = 5;  
x2 - 0.5(5) = -3.5 => x2 = -1  
x1 -1 +5 = 6 => x1 = 2  

### Note that throughout the operation, we only use + / - between rows; if we have to use multiply we just use scaler multiplications