---
title: "20181016a Optimization in differentiation"
author: "cheungngo"
date: "October 16, 2018"
output: word_document
---

```{r,message=FALSE}
library(Ryacas)
library(mosaic)
```


### Critical points

f'(x) > 0 => increasing value  
f'(x) < 0 => decreasing value  
f'(x) = 0 => min / max value  

f''(x) > 0 => trough  
f''(x) < 0 => peak  
f''(x) = 0 => can be an inflexion  

Let f(x) = 2*x^3 - 9*x^2 + 1  
Find the critical points  

```{r}
x = Sym('x')
f = 2*x^3 - 9*x^2 + 1
df = deriv(f,x)
ddf = deriv(df,x)
```

```{r}
# Finding the x of the critical points
Solve(df==0,x)
# Finding the y of the critical points
Eval(f,list(x = c(0,3)))
```

```{r}
# Determining max or min
Eval(ddf,list(x = c(0,3)))
```

Conclusion: (0,1) being maxima; (3,-26) being minima  

```{r}
plotFun(2*x^3 - 9*x^2 + 1~x)
```


### Global or local maxima / minima  

To get the biggest maxima, we just calculate all the f(x) of the critical points  
But the greatest maxima may not necessarily mean it is the greatest among f(x)  
In this way we need to estimate f(x) towards the extremes  

Let f(x) = x^3 - 6*x^2 + 11*x - 6 

```{r}
f = x^3 - 6*x^2 + 11*x - 6
df = deriv(f,x); df = Simplify(df); df
ddf = deriv(df,x); ddf = Simplify(ddf); ddf
```

```{r}
# x in critical points
Solve(df==0,x)
# y in critical points
Eval(f,list(x=c((root(12, 2) + 12)/6, (12 - root(12, 2))/6)))
```

```{r}
# Max or Min
Eval(ddf,list(x=c((root(12, 2) + 12)/6, (12 - root(12, 2))/6)))
```

```{r}
# Global or Local
Limit(f,x,Infinity)
Limit(f,x,-Infinity)
```

So the critical points are only local maxima and minima  

```{r}
# Confirmed by graph
plotFun(x^3 - 6*x^2 + 11*x - 6 ~ x)
```

