---
title: "20181019k application of the Lagrange multiplier"
author: "cheungngo"
date: "19 October 2018"
output: word_document
---

### Principles and worked example

Utility U(x,y) = x1^(1/3)*x2^(1/2)  
Budget constrant M = p1x1 + p2x2  
i.e. L = x1^(1/3)*x2^(1/2) - lambda(p1x1 + p2x2 - M)  

The Lagrange multiplier is said to be equal to du/dM

Therefore, by finding lambda, we find dU/dM, thus able to find the change in utility as the budget changes

```{r,message=FALSE}
library(Ryacas)
library(mosaic)
library(manipulate)
```

```{r}
x1 = Sym('x1'); x2 = Sym('x2'); lambda = Sym('lambda')
p1 = Sym('p1'); p2 = Sym('p2'); M = Sym('M')
U = x1^(1/3)*x2^(1/2) 
L = U - lambda*(p1*x1 + p2*x2 - M)
dLx1 = deriv(L,x1); dLx1
dLx2 = deriv(L,x2); dLx2
dLlambda = deriv(L,lambda); dLlambda
```

```{r}
Solve(dLx1==0,lambda)
Solve(dLx2==0,lambda)
```

```{r}
PrettyForm((x1^-(2/3) * x2^0.5)/p1/3 == x1^(1/3) * (0.5 * x2^-0.5)/p2)
```

=> x2/(3p1) == x1/(2p2)
=> x1 = x2(2/3)(p2/p1)

```{r}
Solve(List(x2/(3*p1) == x1/(2*p2), dLlambda==0),
      List(x1,x2))
```

x2 == 3M/(5 * p2)
x1 == x2(2/3)(p2/p1) == 3*M/5(2/3)(1/p1) == 2M/5p1

```{r}
# getting the lambda
lambdax = makeFun(0.3333333333 * (x1^-0.6666666666 * x2^0.5)/p1 ~ x1&x2&p1)
Simplify(lambdax(x1=2*M/5/p1,x2=3*M/(5 * p2),p1=p1))
PrettyForm(((2 * M/(5 * p1))^-0.6666666666 * (3 * M/(5 * p2))^0.5)/p1/3)
```

```{r}
lambda_LHS = 2^(-2/3)/5^(-2/3)*3^(0.5)/5^(1/2)/3*p1^(2/3-1)*p2^(-1/2)*M^(0.5-2/3)
lambda_LHS
```


```{r}
# Indirect utility function
# U(x1,x2) => U(M,p1,p1) by substitution
Ux = makeFun(x1^(1/3)*x2^(1/2)~x1&x2)
Ux(x1=2*M/5/p1,x2=3*M/(5 * p2))
PrettyForm((2 * M/(5 * p1))^0.333333333333333 * (3 * M/(5 * p2))^0.5)
```

U(M,p1,p2) == 2^(1/3)*3^(1/2)/5^(5/6) p1^(-1/3) p2^(-1/2) M^(1/2+1/3)

```{r}
2^(1/3)*3^(1/2)/5^(5/6)*p1^(-1/3)*p2^(-1/2)*M^(1/2+1/3)
```

```{r}
# Trying to get the derivatives of U
dUM = deriv(0.570727705645511 * (p1^-0.3333333333 * p2^-0.5 * M^0.833333333333333),M)
lambda_RHS = Simplify(dUM)
lambda_LHS; lambda_RHS
```

### Therefore, by finding lambda, we find dU/dM, thus able to find the change in utility as the budget changes