---
title: "20181019b partial fractions"
author: "cheungngo"
date: "19 October 2018"
output: word_document
---

### Partial fractions demonstrated by examples

This is done when the denominator could be factorized  

Find Int x / (x^2 - x - 2) dx  

(Actually the nominator can be many other things other than x)

```{r,message=FALSE}
library(Ryacas)
library(mosaic)
```

```{r}
# LHS
x = Sym('x')
LHS = Integrate(x / (x^2 - x - 2), x)
LHS
```

```{r}
# RHS
# Let f(x) == x and g(x) == (x^2 - x - 2)
# Factorize: x^2 - x - 2 == (x+1)*(x-2)
# It then becomes x / (x+1)*(x-2)
# Consider: A1 / (x+1) + A2 / (x-2) == f(x) / g(x);
# f(x) == A1(x-2) + A2(x+1)
# In this case, f(x) == x => x == A1(x-2) + A2(x+1)
# The cover-up rule: let x = 2  and x = -1
# For x = 2, 2 == A2*3 => A2 = 2/3
# For x = -1, -1 == A1*-3 => A1 = 1/3
```

```{r}
# Consider: A1 / (x+1) + A2 / (x-2) == f(x) / g(x)
# f(x) / g(x) == 1/(3*(x+1)) + 2/(3*(x-2))
Integrate(1/(3*(x+1)) + 2/(3*(x-2)),x)
```

### Summary of the Prinicple

For Int f(x) / g(x) dx;  
While g(x) factorizable into (x-a1)*(x-a2)  
f(x) / g(x) == A1/(x-a1) * A2/(x-a2)  
And f(x) == A1(x-a2) + A2(x-a1)  
A1 and A2 can thus be found by the 'cover-up' rule  
And the A1/(x-a1) * A2/(x-a2) can thus be integrated by substitution